Bayesian Nonparametric System Reliability\\ using Sets of Priors

### 2015-12-20 ijar-npb-ms.tex
work on sec 6 and notation n_k

As a simple example, assume that we expect $\ptk$ to be about $0.75$ for a certain $k$ and $t$,
so we choose $\yktz = 0.75$,
and that we value this choice of mean functioning probability with $\nktz = 8$,
i.e., like having seen $8$ observations with a mean $0.75$.
Assume further that there are $n^k = 16$ components of type $k$ in the test data.
When we observe $s^k_t = 12$ components functioning at time $t$,
so $s^k_t/n^k = 0.75$ as we expect,
then the updated parameters are $\nktn = 24, \yktn=0.75$.
Unexpectedly observing that no component functions at time $t$ instead
leads to parameters $\nktn = 24, \yktn=0.25$.
The prior and the posteriors based on these two observation scenarios
are depicted in Figure~\ref{fig:singleprior-pdc}
through their Beta densities and cdfs (left)
and corresponding Beta-binomial probabilities $P(C^k_t = l_k\mid s^k_t)$ and cmfs
for the case of $m_k = 5$ (right).

@@

\section{Sets of Beta Priors}

As was shown by \citet{2009:WalterAugustin}, %Walter \& Augustin (2009),
we can have both tractability and meaningful reaction to prior-data conflict
by using sets of priors $\MktZ$ produced by parameter sets $\PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu]$. 
(A detailed discussion of different choices for $\PktZ$ is given in \citet[\S 3.1]{2013:diss-gw}.)
In our model, each prior parameter pair $(\nktz, \yktz) \in \PktZ$
corresponds to a Beta prior, thus $\MktZ$ is a set of Beta priors.
The set of posteriors $\MktN$ is obtained by updating each prior in $\MktZ$ according to Bayes' Rule.
This element-by-element updating may seem ad-hoc, but can be rigorously justified
as ensuring coherence \citep[\S ***]{1991:walley}, and was termed ``Generalized Bayes' Rule'' by \citet[\S 6***]{1991:walley}.
Due to conjugacy, $\MktN$ is a set of Beta distributions with parameters $(\nktn, \yktn)$
obtained from updating $(\nktz, \yktz) \in \PktZ$ according to \eqref{eq:nyupdate},
and so can be described efficiently through the set of updated parameters
\begin{align*}
\PktN &= \Big\{ (\nktn, \yktn) \mid (\nktz, \yktz) \in \PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu] \Big\}\,.
\end{align*}
Such rectangular prior parameter sets $\PktZ$ have been shown
to balance desirable model properties and ease of elicitation quite well.
%
For each component type $k$ and time point $t$,
one has to choose only the four parameters $\nktzl, \nktzu, \yktzl, \yktzu$
(so $4 \times |{\cal T}|$ parameters are needed to define the set of prior distributions characterizing
prior knowledge on the survival function of components of type $k$);
notwithstanding this, the model has favourable inference properties.

Central to these inference properties is that
while the prior parameter set $\PktZ$ is rectangular,
the posterior parameter set $\PktN$ is not rectangular anymore;
indeed, the shape of $\PktN$ depends on the presence of prior-data conflict,
which is naturally operationalized as $s^k_t/n^k_t \not\in [\yktzl, \yktzu]$:
prior-data conflict is present when the observed fraction of functioning components
is outside its expected range.

@@

In case of prior-data conflict ($s^k_t/n^k_t \not\in [\yktzl, \yktzu]$),
$\PktN$ has instead the so-called banana shape,
arising from $\yktn$ intervals being shifted closer towards $s^k_t/n^k_t$
for lower $\nktn$ values than for higher $\nktn$ values, see Figure***.
Overall, this results in a wider $\yktn$ interval as compared to the non-conflict case, 
reflecting the extra uncertainty due to prior-data conflict,
thus making more cautious probability statements about $p_t^k$.

@@

As mentioned in Section~\ref{sec:nonparamapproach}, for the actual functioning probabilities $\ptk$
it should hold that $p^k_{t_j} \ge p^k_{t_{j+1}}$. ***or $p^k_{t_i} \ge p^k_{t_i}$ for all $i < j$.
%***not necessary to ensure $p^k_{t_j} \ge p^k_{t_{j+1}}$ via $\yzl_{k,t_j} \ge \yzu_{k,t_{j+1}}$
%as for relatively dense $t$ grids, neighbouring intervals should be able to overlap***
%(Prior guess for $p^k_t$ interpretation would say that $\yktz$ cannot increase,
%but does this need to hold for the prior guess range bound $\yktzu$?)***
This naturally translates to conditions on the prior guess for $\ptk$,
such that we propose that 
$\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$ should hold.
Due to $s^k_t/n_k$ decrasing in $t$ and the weighted average property of the update step for $\ykt$,
this ensures that 
$\ynu_{k,t_j} \ge \ynu_{k,t_{j+1}}$ and $\ynl_{k,t_j} \ge \ynl_{k,t_{j+1}}$.
In a situation where we are quite sure about the functioning probability for low $t$,
but are unsure about what happens for larger $t$ and so want to be maximally cautious for those $t$,
then we can let $\yktzl$ drop to (almost) 0, but $\yktzu$ should not increase.
%***Require $\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$.
%If we know, e.g., pretty much what's going on for low $t$
%but want to be much more cautious for high $t$,
%then $\yktzl$ can drop to 0, but $\yktzu$ cannot increase.

We think it is not advisable to express certainty about expected functioning probabilities
with $\nktz$ bounds that vary strongly over $t$.
%***Better to express certainty of prior guess via $\nktz$ bounds?\\
With (strongly) differing $\nktz$ bounds, monotonicty for $\yktn$ bounds cannot be ensured.
%it can happen that, e.g., $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$!\\
E.g., when $\yzu_{k,t_j} = \yzu_{k,t_{j+1}}$, $\yzl_{k,t_j} = \yzl_{k,t_{j+1}}$,
and $s^k_{t_j}/n_k \in [\yzl_{k,t_j}, \yzu_{k,t_j}]$ (so there is no prior-data conflict),
and furthermore no failures happening in $[t_j, t_{j+1}]$, i.e., $s^k_{t_{j+1}}/n_k = s^k_{t_{j}}/n_k$, then
\begin{itemize}
\item for $\nzu_{k,t_j} < \nzu_{k,t_{j+1}}$ we get $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$, and
\item for $\nzu_{k,t_j} > \nzu_{k,t_{j+1}}$ we get $\ynl_{k,t_j} < \ynl_{k,t_{j+1}}$.
\end{itemize}
Again, this follows from the weighted average property of $\yktn$;
for constant $\yktz$, the lower $\nktz$ the closer $\yktn$ is to $s^k_t/n_k$.
% one can think of the update between interval and precise value: the lower $n$, the 'faster' the interval will shrink!)\\
Therefore, we propose to take the same $\nktz$ bounds for all $t$, or, at most, to have it change with $t$ gradually only.

%***generally require $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$
%to get proper posteriors***
Generally, one should choose $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.

***what to do in extreme tail of component distributions?
How sensitive is the method to choosing prior (lower, upper) mean near-zero (how near?) for $p^k_t$ at high $t$'s?***

***elicitation: we should also consider the case where $\yktz$ bounds are not given for all $t \in {\cal T}$,
as it will be too time-consuming to elicit $\yktz$ bounds for all $t$ in a dense grid.
(Calculating $\Rsys(t)$ for a too coarse grid only will waste information from data.)
Idea: ``fill up the $t$ grid at all $t$'' with least-committal bounds, i.e.: 
at all $t$ not elicited, take $\yktzu$ equal to last (in time sequence) elicited $\yktzu$,
and $\yktzl$ equal to next (in time sequence) elicited $\yktzl$.

***possible elicitation procedure: starting with $\yktz$ bounds for `middle' or `typical' $t$,
then $\yktz$ bounds for $t$ smaller than `middle' $t$ (or halfway between 0 and `middle' $t$?) and for $t$ larger than `middle' $t$,
refine further as possible.

@@

***code published in Louis' \textbf{R} package \texttt{ReliabilityTheory}
(which contains the function to calculate the survival signature)***

>>>>

As a simple example, assume that we expect $\ptk$ to be about $0.75$ for a certain $k$ and $t$,
so we choose $\yktz = 0.75$,
and that we value this choice of mean functioning probability with $\nktz = 8$,
i.e., like having seen $8$ observations with a mean $0.75$.
Assume further that there are $n_k = 16$ components of type $k$ in the test data.
When we observe $s^k_t = 12$ components functioning at time $t$,
so $s^k_t/n_k = 0.75$ as we expect,
then the updated parameters are $\nktn = 24, \yktn=0.75$.
Unexpectedly observing that no component functions at time $t$ instead
leads to parameters $\nktn = 24, \yktn=0.25$.
The prior and the posteriors based on these two observation scenarios
are depicted in Figure~\ref{fig:singleprior-pdc}
through their Beta densities and cdfs (left)
and corresponding Beta-binomial probabilities $P(C^k_t = l_k\mid s^k_t)$ and cmfs
for the case of $m_k = 5$ (right).

@@

\section{Sets of Beta Priors}
\label{sec:setsofbetapriors}

As was shown by \citet{2009:WalterAugustin}, %Walter \& Augustin (2009),
we can have both tractability and meaningful reaction to prior-data conflict
by using sets of priors $\MktZ$ produced by parameter sets $\PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu]$. 
(A detailed discussion of different choices for $\PktZ$ is given in \citet[\S 3.1]{2013:diss-gw}.)
In our model, each prior parameter pair $(\nktz, \yktz) \in \PktZ$
corresponds to a Beta prior, thus $\MktZ$ is a set of Beta priors.
The set of posteriors $\MktN$ is obtained by updating each prior in $\MktZ$ according to Bayes' Rule.
This element-by-element updating may seem ad-hoc, but can be rigorously justified
as ensuring coherence \citep[\S 2.5]{1991:walley}, and was termed ``Generalized Bayes' Rule'' by \citet[\S 6.4]{1991:walley}.
Due to conjugacy, $\MktN$ is a set of Beta distributions with parameters $(\nktn, \yktn)$
obtained from updating $(\nktz, \yktz) \in \PktZ$ according to \eqref{eq:nyupdate},
and so can be described efficiently through the set of updated parameters
\begin{align*}
\PktN &= \Big\{ (\nktn, \yktn) \mid (\nktz, \yktz) \in \PktZ = [\nktzl, \nktzu] \times [\yktzl, \yktzu] \Big\}\,.
\end{align*}
Such rectangular prior parameter sets $\PktZ$ have been shown
to balance desirable model properties and ease of elicitation quite well.
%
For each component type $k$ and time point $t$,
one has to choose only the four parameters $\nktzl, \nktzu, \yktzl, \yktzu$
(so $4 \times |{\cal T}|$ parameters are needed to define the set of prior distributions characterizing
prior knowledge on the survival function of components of type $k$);
notwithstanding this, the model has favourable inference properties.

Central to these inference properties is that
while the prior parameter set $\PktZ$ is rectangular,
the posterior parameter set $\PktN$ is not rectangular anymore;
indeed, the shape of $\PktN$ depends on the presence of prior-data conflict,
which is naturally operationalized as $s^k_t/n_k \not\in [\yktzl, \yktzu]$:
prior-data conflict is present when, at time $t$, the observed fraction of functioning components
is outside its expected range.

@@

In case of prior-data conflict ($s^k_t/n_k \not\in [\yktzl, \yktzu]$),
$\PktN$ has instead the so-called banana shape,
arising from $\yktn$ intervals being shifted closer towards $s^k_t/n_k$
for lower $\nktn$ values than for higher $\nktn$ values, see Figure***.
Overall, this results in a wider $\yktn$ interval as compared to the non-conflict case, 
reflecting the extra uncertainty due to prior-data conflict,
thus making more cautious probability statements about $p_t^k$.

@@

As mentioned in Section~\ref{sec:nonparamapproach}, for the actual functioning probabilities $\ptk$
it should hold that $p^k_{t_j} \ge p^k_{t_{j+1}}$. (***or $p^k_{t_i} \ge p^k_{t_i}$ for all $i < j$).
%***not necessary to ensure $p^k_{t_j} \ge p^k_{t_{j+1}}$ via $\yzl_{k,t_j} \ge \yzu_{k,t_{j+1}}$
%as for relatively dense $t$ grids, neighbouring intervals should be able to overlap***
%(Prior guess for $p^k_t$ interpretation would say that $\yktz$ cannot increase,
%but does this need to hold for the prior guess range bound $\yktzu$?)***
This naturally translates to conditions on the prior guess for $\ptk$,
such that we propose that 
$\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$ should hold.
Due to $s^k_t/n_k$ decrasing in $t$ and the weighted average property of the update step for $\ykt$,
this ensures that 
$\ynu_{k,t_j} \ge \ynu_{k,t_{j+1}}$ and $\ynl_{k,t_j} \ge \ynl_{k,t_{j+1}}$.
In a situation where we are quite sure about the functioning probability for low $t$,
but are unsure about what happens for larger $t$ and so want to be maximally cautious for those $t$,
then we can let $\yktzl$ drop to (almost) 0, but $\yktzu$ should not increase.
%***Require $\yzu_{k,t_j} \ge \yzu_{k,t_{j+1}}$ and $\yzl_{k,t_j} \ge \yzl_{k,t_{j+1}}$.
%If we know, e.g., pretty much what's going on for low $t$
%but want to be much more cautious for high $t$,
%then $\yktzl$ can drop to 0, but $\yktzu$ cannot increase.

We think it is not advisable to express certainty about expected functioning probabilities
with $\nktz$ bounds that vary strongly over $t$.
%***Better to express certainty of prior guess via $\nktz$ bounds?\\
With (strongly) differing $\nktz$ bounds, monotonicty for $\yktn$ bounds cannot be ensured.
%it can happen that, e.g., $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$!\\
E.g., when $\yzu_{k,t_j} = \yzu_{k,t_{j+1}}$, $\yzl_{k,t_j} = \yzl_{k,t_{j+1}}$,
and $s^k_{t_j}/n_k \in [\yzl_{k,t_j}, \yzu_{k,t_j}]$ (so there is no prior-data conflict),
and furthermore no failures happening in $[t_j, t_{j+1}]$, i.e., $s^k_{t_{j+1}}/n_k = s^k_{t_{j}}/n_k$, then
\begin{itemize}
\item for $\nzu_{k,t_j} < \nzu_{k,t_{j+1}}$ we get $\ynu_{k,t_j} < \ynu_{k,t_{j+1}}$, and
\item for $\nzu_{k,t_j} > \nzu_{k,t_{j+1}}$ we get $\ynl_{k,t_j} < \ynl_{k,t_{j+1}}$.
\end{itemize}
Again, this follows from the weighted average property of $\yktn$;
for constant $\yktz$, the lower $\nktz$ the closer $\yktn$ is to $s^k_t/n_k$.
% one can think of the update between interval and precise value: the lower $n$, the 'faster' the interval will shrink!)\\
It is possible to construct similar examples also with regard to the lower bound $\nzl_{k,t_j}$.
Therefore, we propose to take the same $\nktz$ bounds for all $t$, or, at most, to have it change with $t$ gradually only.

Generally, the interpretation as pseudocount or prior strength should guide the choice of bounds for $\nktz$; 
low values for $\nktz$ as compared to the test sample size $n_k$
give low weight to the prior expected functioning probability intervals $[\yktzl, \yktzu]$,
and the location of posterior intervals $[\yktnl, \yktnu]$ will be dominated by the location of $s^k_t/n_k$.
Furthermore, the length of $[\yktnl, \yktnu]$ is shorter for low $\nktz$ values;
in a no-conflict situation, when $\nktzu = n_k$ $[\yktnl, \yktnu]$ has half the length of $[\yktzl, \yktzu]$.
In contrast, high values for $\nktz$ will lead to slower learning and wider $\yktn$ intervals,
which means more cautious posterior inferences.
The difference betweeen $\nktzu$ and $\nktzl$ determines the strength of the prior-data conflict sensitivity;
as is clear from Figure~***, the wider the $\nktz$ interval,
the wider $[\yktnl, \yktnu]$ in case of conflict.
So it seems useful to choose $\nktzl = 1$ or $\nktzl = 2$,
while choosing $\nktzu$ with help of the half-width rule as described above.

%***generally require $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$
%to get proper posteriors***
As mentioned in Section~\ref{sec:setsofbetapriors},
it is not advisable to choose $\yktzl = 0$ and $\yktzu = 1$.
For any $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$,
this can lead to improper posterior predictive distributions if $l_k=0$ or $l_k=m_k$.
%Generally, one should choose $\yktzl > 0$ and $\yktzu < 1$ for all $t \not\in \big(\min(\vec{t}^k), \max(\vec{t}^k)\big)$.
But it is possible to choose values close to $0$ and $1$, respectively,
and due to the linear update step for $\yktn$ (see Eq.~\eqref{eq:nyupdate}),
posterior inferences are not overly sensitive to whether $\yktnu = 0.99$ or $\yktnu = 0.9999$.
%***what to do in extreme tail of component distributions?
%How sensitive is the method to choosing prior (lower, upper) mean near-zero (how near?) for $p^k_t$ at high $t$'s?***
Likewise, our nonparametric method does not bring unintuitive tail behaviour
as some parametric methods do;
there is no problem in, e.g., assigning $\yktnu$ near-zero for high $t$ if prior knowledge suggests so.

%***elicitation: we should also consider the case where $\yktz$ bounds are not given for all $t \in {\cal T}$,
%as it will be too time-consuming to elicit $\yktz$ bounds for all $t$ in a dense grid.
While it is possible to set the bounds $\yktzl$ and $\yktzu$ for each $t \in {\cal T}$ individually,
in practice this will be often too time-consuming when ${\cal T}$ forms a dense grid.
%(Calculating $\Rsys(t)$ for a too coarse grid only will waste information from data.)
Switching to a coarser time grid will waste information from data,
as then failure times in the test data are rounded up to the next $t \in {\cal T}$.
We rather propose to elicit bounds only for a subset of ${\cal T}$,
and to fill up the $t$ grid with the least committal bounds, i.e.,
taking $\yktzu$ equal to last (in the time sequence) elicited $\yktzu$,
an likewise $\yktzl$ equal to next (in time sequence) elicited $\yktzl$.
A possible elicitation procedure in this vein could be
to start with eliciting $\yktz$ bounds for a few `central' time points $t$,
filling up the grid as described above accordingly,
and then to further refine the obtained bounds if necessary.
%`middle' or `typical' $t$,
%then $\yktz$ bounds for $t$ smaller than `middle' $t$ (or halfway between 0 and `middle' $t$?) and for $t$ larger than `middle' $t$,

@@

***code published in Louis' \textbf{R} package \texttt{ReliabilityTheory},
function \texttt{nonParBayesSystemInferencePriorSets()};
the package also contains the function \texttt{computeSystemSurvivalSignature()}
to calculate the survival signature for a given system layout (reliability block diagramm)***
